---
title: "MA678 Homework 5"
author: "Jing Xu"
date: "10/25/2022"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load("rstanarm","dplyr","MASS","haven","ggplot2","censReg","brms", "glmx")
remotes::install_github("avehtari/ROS-Examples", subdir = "rpackage")
```

## 15.1 Poisson and negative binomial regression
The folder `RiskyBehavior` contains data from a randomized trial targeting couples at high risk of HIV infection. The intervention provided counseling sessions regarding practices that could reduce their likelihood of contracting HIV. Couples were randomized either to a control group, a group in which just the woman participated, or a group in which both members of the couple participated. One of the outcomes examined after three months was "number of unprotected sex acts."  

### a) 
Model this outcome as a function of treatment assignment using a Poisson regression. Does the model fit well? Is there evidence of overdispersion?  

```{r}
library(rosdata)
data(risky)
risky$fupacts <- round(risky$fupacts)
risky$women_alone <- as.factor(risky$women_alone)
risky$couples <- as.factor(risky$couples)
fit_1a <- glm(fupacts ~ couples +women_alone, family = poisson(link = "log"), data = risky)
summary(fit_1a)
#check for overdispersion
fit_1a$deviance/fit_1a$df.residual<=1
#check for fitting
pchisq(fit_1a$deviance, fit_1a$df.residual, lower.tail = F)
```

### b) 
Next extend the model to include pre-treatment measures of the outcome and the additional pre-treatment variables included in the dataset. Does the model fit well? Is there evidence of overdispersion?  

```{r}
fit_1b <- glm(fupacts ~ log(bupacts +1) + sex + couples + women_alone + bs_hiv, family = poisson(link = "log"), data = risky)
summary(fit_1b)
#check for overdispersion
fit_1b$deviance/fit_1b$df.residual<=1
#check for fitting
pchisq(fit_1b$deviance, fit_1b$df.residual, lower.tail = F)
```

### c) 
Fit a negative binomial (overdispersed Poisson) model. What do you conclude regarding effectiveness of the intervention?

```{r}
fit_1c <- glm.nb(fupacts ~ log(bupacts +1) + sex + couples + women_alone + bs_hiv, data = risky)
summary(fit_1c)
#check for overdispersion
fit_1c$deviance/fit_1c$df.residual<=1
#check for fitting
pchisq(fit_1c$deviance, fit_1c$df.residual, lower.tail = F)
```

### d) 
These data include responses from both men and women from the participating couples. Does this give you any concern with regard to our modeling assumptions? 

```{r}
```
there's a linear relationship between these two variables, so the models don't fit well

## 15.3 Binomial regression
Redo the basketball shooting example on page 270, making some changes:  

### (a) 
Instead of having each player shoot 20 times, let the number of shots per player vary, drawn from the uniform distribution between 10 and 30.  
```{r}
set.seed(100)
N <- 100
height <- rnorm(N, 72, 3)
p <- 0.4 + 0.1*(height - 72)/3
n <- floor(runif(N, min=10, max=31))
y <- rbinom(N, n, p)
data <- data.frame(n = n, y = y, height = height)
```

### (b) 
Instead of having the true probability of success be linear, have the true probability be a logistic function, set so that Pr(success) = 0.3 for a player who is 5'9" and 0.4 for a 6' tall player. 

```{r}
b1 <- (logit(.4) - logit(.3))/3
b0 <- (logit(.4) - 72*b1)
set.seed(100)
N <- 100
height <- rnorm(N, 72, 3)
p <- invlogit(b0 + b1*height)
n <- floor(runif(N, min=10, max=31))
y <- rbinom(N, n, p)
data <- data.frame(n = n, y = y, height = height)
```


## 15.7 Tobit model for mixed discrete/continuous data
Experimental data from the National Supported  Work example are in the folder `Lalonde`. Use the treatment indicator and pre-treatment variables to predict post-treatment (1978) earnings using a Tobit model. Interpret the model coefficients. 

```{r}
lalonde <- haven::read_dta("http://www.nber.org/~rdehejia/data/nsw_dw.dta")
fit_7 <- censReg(formula = re78 ~ re75 + re74, data = lalonde)
summary(fit_7)
```


## 15.8 Robust linear regression using the t model
The folder `Congress` has the votes for the Democratic and Republican candidates in each U.S. congressional district in 1988, along with the parties' vote proportions in 1986 and an indicator for whether the incumbent was running for reelection in 1988. For your analysis, just use the elections that were contested by both parties in both years.  

```{r}
library(rosdata)
data(congress)
```

### (a) 
Fit a linear regression using `stan_glm` with the usual normal-distribution model for the errors predicting 1988 Democratic vote share from the other variables and assess model fit.

```{r}
df <- data.frame(vote = congress$v88_adj, past_vote = congress$v86_adj, inc = congress$inc88)
fit_8a <- stan_glm(vote ~ past_vote + inc, data = df, refresh = 0)
print(fit_8a)
pp_check(fit_8a)

```

### (b) 
Fit the same sort of model using the `brms` package with a $t$ distribution, using the `brm` function with the student family. Again assess model fit.  

```{r}
fit_8b <- brm(vote ~ past_vote + inc, data = df, family = student(link = "identity"), refresh = 0)
```

### (c) 
Which model do you prefer? 


## 15.9 Robust regression for binary data using the robit model
Use the same data as the previous example with the goal instead of predicting for each district whether it was won by the Democratic or Republican candidate.  

### (a) 
Fit a standard logistic or probit regression and assess model fit.

```{r}
fit_9a <- stan_glm(v88_adj ~ v86_adj + inc86, family = binomial(link = "probit"), data = congress, refresh = 0) 
summary(fit_9a)
pp_check(fit_9a)
```

### (b) 
Fit a robit regression and assess model fit.

```{r}
fit_9b <- rlm(v88_adj ~ v86_adj + inc86, family = binomial(link = "probit"), data = congress) 
summary(fit_9b)
plot(fit_9b, which = 1)
```

### (c) 
Which model do you prefer? 
robit regression model provides better residuals for conparison.

## 15.14 Model checking for count data
The folder `RiskyBehavior` contains data from a study of behavior of couples at risk for HIV; see Exercise 15.1. 

### (a) 
Fit a Poisson regression predicting number of unprotected sex acts from baseline HIV status. Perform predictive simulation to generate 1000 datasets and record the percentage of observations that are equal to 0 and the percentage that are greater than 10 (the third quartile in the observed data) for each. Compare these to the observed value in the original data.

```{r}
risky$bs_hiv <- as.factor(risky$bs_hiv)
fit_15a <- glm(fupacts ~ bs_hiv, family = poisson, data = risky)

```

### (b) 
Repeat (a) using a negative binomial (overdispersed Poisson) regression.

```{r}
```

### (c) 
Repeat (b), also including ethnicity and baseline number of unprotected sex acts as inputs.

```{r}
```


## 15.15 Summarizing inferences and predictions using simulation
Exercise 15.7 used a Tobit model to fit a regression with an outcome that had mixed discrete and continuous data. In this exercise you will revisit these data and build a two-step model: 
(1) logistic regression for zero earnings versus positive earnings, and 
(2) linear regression for level of earnings given earnings are positive. 
Compare predictions that result from each of these models with each other. 

```{r}
lalonde <- haven::read_dta("http://www.nber.org/~rdehejia/data/nsw_dw.dta")
data_1 <- lalonde
data_1$re78 <- ifelse((data_1$re78 > 0), 1, 0)
fit15_1 <- glm(re78 ~ treat + sample + age + education + black + hispanic + married + nodegree, family = binomial(link = "log"), data = data_1)
print(fit15_1)
data_2 <- lalonde[which(lalonde$re78 > 0), ]
fit15_2 <- glm(re78 ~ treat + sample + age + education + black + hispanic + married + nodegree, data = data_2)
print(fit15_2)
```
